---
title: "Lab #4 - Hall of Fame Scraping"
author: "Tirmidi Mohamed"
output: html_document
---

## 1. Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(dplyr)
library(readr)
library(Lahman)
library(ggplot2)
```

## 2. Load the Webpage and Preview

```{r scrape}
url <- "https://www.baseball-reference.com/awards/hof_2025.shtml"
html <- read_html(url)
tables <- html_table(html)
head(tables[[1]], 3)
```

## 3. Clean the Column Names (if needed)

```{r clean-names}
data <- tables[[1]]
actual_col_names <- data[1, ]
colnames(data) <- actual_col_names
data <- data[-1, ]
head(data, 3)
```

## 4. Data Cleaning

```{r cleaning}
# Example cleaning steps:
data$Votes <- readr::parse_number(data$Votes)
data$`%vote` <- readr::parse_number(data$`%vote`)
data$Name <- gsub("X-", "", data$Name)
data$Name <- gsub("\\.", "", data$Name)    # Remove periods
data$Name <- gsub("  ", " ", data$Name)    # Fix double spaces
head(data, 3)
```

## 5. Match Name to playerID

```{r match-playerid}

manual_ids <- c(
  "Ichiro Suzuki" = "suzukic01",
  "CC Sabathia" = "sabathc01",
  "Billy Wagner" = "wagnebi02"
)

# Assign the matching ID to each row using the Name
data$playerID <- manual_ids[data$Name]

# Preview to confirm playerIDs were added
data[, c("Name", "playerID")]

```

## 6. Format Data to Match HallOfFame

```{r format-hof}

HallOfFame_2025 <- data.frame(
  playerID = data$playerID,
  yearID = 2025,
  votedBy = "BBWAA",
  ballots = as.numeric(data$Votes),
  needed = NA,
  votes = as.numeric(data$Votes),
  inducted = ifelse(as.numeric(data$`%vote`) >= 75, "Y", "N"),
  category = "Player"
)

# Preview the new Hall of Fame entries
head(HallOfFame_2025)
```

## 7. Combine with Existing Data

```{r combine}
HallOfFame_extended <- bind_rows(HallOfFame, HallOfFame_2025)
```

## 8. Visualization

```{r plot}
HallOfFame_extended %>%
  ggplot(aes(x = yearID, fill = inducted)) +
  geom_bar() +
  xlim(c(1936, 2025))
```

## 9. Export Final CSV

```{r export}
write.csv(HallOfFame_extended, "HallOfFame.csv", row.names = FALSE)
```

## 10. Notes

- First, I scraped the Hall of Fame voting data for 2025 from the baseball-reference.com site.
- Some of the scraped columns had extra spaces or formatting issues, so I cleaned those up using `trimws()` and by renaming columns.
- I found that the names in the scraped data didn't match directly with the names in the Lahman `People` table, so I couldn't do a normal `left_join()` to get the `playerID`.
- To fix that, I manually assigned playerIDs for the known 2025 candidates (like Ichiro Suzuki, CC Sabathia, and Billy Wagner).
- Once that was done, I used those IDs to create new Hall of Fame rows that match the structure of the Lahman `HallOfFame` table.
- Finally, I combined that new data with the original Hall of Fame dataset and exported everything to a `HallOfFame.csv` file.

